### 2018-09-19 draft.tex
Add Docker Swarm software description

\documentclass[10pt]{book}

\usepackage{todonotes}

\author{Polonio Davide}
\title{A thesis draft}
% The outline'd be something like this:
%  Thesis outline
% 
% - Introduction about the VIBES project (context, 5G, the need for a new 
% network architecture, proposed solution, requisites)
%   + Related works
% - Analysis of the available technologies (kubernetes, openstack, openvswitch, 
% Docker, Docker swarm)
%   + Choseen tecnhologies and description of the 'reason why'
% - Project description
%   + Initial solution with Openstack and Tacker
%   + From Tacker to Openbaton
%   + From Openbaton with Openstack to Openbaton with Docker
%   + Exploration of Docker orchestrators (Swam vs Kubernetes)
%   + Kubernetes internal composition (GlusterFS, Ingress, Harbor, metrics)
%   + Openbaton and Kubernetes integration (lack of a implemented Driver, 
% integration with existinf VNFM)
% - Case study
%   + Analysis of TCP/UDP performance
% - Results
% - Future works
% - Conclusions

\makeindex
\begin{document}
 \maketitle

 % This is just a draft, nothing suitable for the final product: sources are in 
 % form of a footnote

 % --- INTRODUCTION ---
\chapter{Introduction}

 Through the history of internet, many factor changed and new innovations came,
but at the end of the day a foundamental key remained: TCP/IP performance are, 
indeed, what the users unconsciously want.\todo{Idea: talk about how TCP/IP 
performes quite well in wired connections, while in wireless one, such as 
wireless/satellite, sucks ass} In this context network optimization plays an 
important role, and it is for this reason that network packets are processed 
while traversing the network in order to make their transmission as fast as 
possible. On top of that, mobile traffic is continuously surging every year, 
with the backbones having more and more data to process. In this context, 
in-hardware VNF solutions are becoming less efficent. Their time-to-deploy, in 
fact, is very high compared to the requirements of the network. With this 
trend, in-hardware solutions will not be able to keep up. Is with this 
considerations that software solutions are being proposed to become the new way 
to deploy network components. While less efficent, they are faster to bring them 
up, offering an unprecedented flexibility\footnote{SDN/NFV-Based Mobile Packet 
Core Network Architectures: A Survey}. Furthermore, new virtualization 
technologies such as Cloud computing and Docker are becoming more and more 
mature for being employed in large-scale production environments, where new 
frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm), allow to 
easily manage the application lifecyle, i.e. deployment, scale up/down
and finally, removal.

\section{The VIBES project}

 This thesis is part of the VIBES project\todo{Talk about VIBES project, add 
some reference, explain what it is.}, where the necessity for better TCP/IP 
transmission through satellite connections is the main requirement. To 
reach this goal, the project specifications suggest to exploit the 5G incoming 
technology and use the NFV-MANO architecture to perform first packets 
elaboration and performance improvment and finally TCP/IP satellite chunk
optimization with the Performance Enhancing Proxy (PEP). The VIBES project 
proposed five technical requirements:
\begin{enumerate}
 \item Analysis of the applicability of current and new Internet protocols in
the proposed VNF-PEP architecture
 \item Implementation of a VNF-PEP prototype
 \item Building of a PoC test platform
 \item VNF-PEP validation and performance tests on 5G use-cases
 \item Demonstration testbed management
\end{enumerate}

\subsection{VNF-PEP architecture and internet protocols}

The analysis of this topic revealed to be trivial: since internet has many 
different protocols that would become infeasible to support all of them at the 
same time, packet incapsulation revealed to be the solution: every packet 
incoming in the VNFs have been previously incapsulated by the MANO, making the 
whole architecture indipendent from the protocol a particolar data-flux uses. 
To achieve this, TUN/TAP\todo{Find out acronym} interfaces were used to deploy 
a tunnel between VNFs, allowing them to pass packets encapsulated in UDP ones.

\subsection{VNF-PEP prototype}

Since our goal shifted into creating a MANO testbed, our prototype doesn't 
include PEP. The VNF architecture was shaped following the container 
orchestrator we decided to use: Kubernetes.

\subsubsection{About kubernetes}

As already introduced before, Kubernetes is an open source software framework 
specialized in container management and orchestration. Developed by Google in 
??\todo{Find out the actual date} and written in Go, it allows different 
machines (called nodes from now on) to create an abstraction layer and to form 
a cluster where is possible to run Docker containers without brothering about 
hardware constraints. Nodes can have specific roles and specialization, all of 
them managed by one or many master nodes, that actually manage the 
orchestration.
On top of that, Kubernetes is virtualization-agnostic, meaning that it offers 
the possibility to change the virtualization software in use (i.e. from 
Container-based technology to Virtual machine software).

\subsubsection{About Docker}

In recent years, with the new hardware capabilities and the recent development 
of in-kernel virtualization systems (such as Hyper-V) this technology 
begin to be adopted. Virtualization allows to run different systems on the 
same machine, making them completely isolated and then more resilient to 
failures. The back of the medal, though, is that virtual machines require large 
amount of resources, especially memory, because solutions like copy-and-write 
of in-memory pages are not viable anymore (since the kernel gets duplicated 
too), leading to duplication of loaded libraries and assets in the memory of 
the host. 
For large deployment containing only simple services (e.g. a backend 
application serving a website or a database) this ends up in a waste of 
resources. \todo{Talk about lxc containers and how Docker solved magically the 
problem.} It's here that, in ??\todo{Find Docker date of birth}, Docker was 
created, basing its solution on an already existing product: Linux Container 
(LXC). From this framework, Docker built an entire ecosystem, consisting in a 
client/server model giving the possibility for users to simply launch, scale and 
delete containers (locally or remotely with Docker-machine), a repository 
system where images, a layer-based ``core'' where containers start running from 
when they are launched, can be stored and tagged in a sort of version control 
system. Finally, another couple of solutions, called Docker-compose, and Docker 
Swarm have the purpose to, respectively, orchestrate containers in a single or 
a clustered system.

\paragraph{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
seen as one logical unit. This layer completely makes the underlying
infrastructure transparent: data management, as network one, are totaly managed
by Docker. Recently, Docker-compose support has been introduced, making
this one-node tool available for clustered systems. At the end of the day, on
one hand Docker Swarm provides an easy way to set up a cluster system with all
the tools configured out-of-the-box, without the necessity to set networking
configurations or installing ad-hoc storage solutions. On the other hand,
though, it doesn't have all the personalization options that Kuberentes offers,
thus making the product less flexible. This foundamental characteristic,
although, makes the two solutions have different use cases leading to different
market needs.
\vspace{0.5cm}

Starting with the first step of creating a MANO able to process incoming data 
packets through VNF functions, we encountered that many networking tools already 
present in the market required some tweaking and some integrations, shifting our 
goal to create a complete  European Telecommunications Standards Institutes 
(ETSI) Management and orchestrator (MANO) testbed instead, following the 
specifications suggested in the RFC 7665, thus implementing only the first three 
requisites, without digging in the satellite data flow optimization.

\subsubsection{Other technologies}\todo{Talk somehow of the other tecnhologies}
\end{document}