### 2018-10-14 res/sections/03-Inscription.tex
Applied thesis template

%**************************************************************
% Dedica
%**************************************************************
\cleardoublepage
\phantomsection
\thispagestyle{empty}
\pdfbookmark{Dedica}{Dedica}

\vspace*{3cm}

\begin{flushright}{
	\slshape
	dedica} \\
	\medskip
\end{flushright}

\medskip