### 2018-10-14 res/sections/04-Credits.tex

%**************************************************************
% Ringraziamenti
%**************************************************************
\cleardoublepage
\phantomsection
\thispagestyle{empty}
\pdfbookmark{Credits}{credits}

%\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\section*{Credits}

\noindent \textit{Ringraziamenti}\\
\bigskip

\noindent\textit{\myLocation, \myTime}
\hfill \myName

\endgroup