### 2018-10-14 res/sections/05-Abstract.tex

%**************************************************************
% Abstract
%**************************************************************
\cleardoublepage
\phantomsection
\section*{Abstract}
\thispagestyle{empty}

\bigskip