### 2018-10-19 res/sections/20-VNFImplementaton.tex
Add new file, add content in background section

\chapter{VNF and Network Service implementation} 
\label{chap:vnf_ns_impl}