### 2018-11-21 res/sections/08-Background.tex
Splitting

\chapter{Background}
\label{chap:background}

In this chapter we are going to explain the most important concepts of the 
upcoming 5G standard, talking about the Virtual Network Functions (VNFs) and 
how first experiments with this new technology already begun. Finally we are 
going to briefly introduce and explain the technologies involved and their role.

\section{The upcoming connectivity standard: 5G}
The continuous innovation in the mobile network connectivity is leading to the
creation of a new standard, the 5G, which is estimated to arrive to the market
consumer in 2020~\cite{iwamura2015ngmn}. Lead by the Next Generation Mobile
Network (NGMN) alliance, a group composed by the major players in the field of
mobile connectivity, the 5G aims to offer not only at the end-user a new way to
browse the web, download and watch interactive content but also to create an
ad-hoc solution for Machine-to-Machine (M2M) data traffic, which is increasing
more than ever thanks to the spreading of IoT devices, which sensors need to 
continuously send data to servers/data-centers. Relatively to \emph{Long Term
Evolution} (LTE), 5G points to have data rates $10$ times better, with $10$
times smaller end-to-end latency and an increased connection density by $100$ 
times~\cite{alliance20155g}.

\begin{table}[t]
\centering
\resizebox{\textwidth}{!}{%
  \begin{tabular}{p{4,5cm}|p{5,5cm}|p{5cm}}
\textbf{Attribute}                                                    & 
\textbf{LTE capability}

                                                                           & 
\textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
\textbf{Data rate (per user)}                                         & Up to 
100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).

                                                                     & 10X 
expected on average and peak rates and 100X expected on cell edge \\
\textbf{End-to-end latency}                                           & 10 ms 
for two-way RAN (pre-scheduled). Typically, up to 50 ms end-to-end if other 
factors are considered (e.g., transmission, CN, internet, proxy servers). & 10X 
(smaller)                                                         \\
\textbf{Mobility}                                                     & 
Functional up to 350 km/h (for certain bands up to 500 km/h). No support for 
civil aviation.                                                                
& 
1.5X                                                                  \\
\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & 
Typically $\sim$2,000 active users/km2.

                                                                           & 
100X                                                                 
\end{tabular}%
}
\caption[5G improvement over LTE]{An extract from the official 5G white paper 
illustrating the improvements of 5G relatively to LTE connections.}
\label{chap:intro:table:ltevs5g}
\end{table}

Part of the new requirements can be satisfied using a large radio spectrum with
higher frequencies. The utilization of higher frequencies, though, mean that the
radio signals can be easily disrupted by physical objects, like buildings and
many geographical elements (such as hills and mountains), clashing with the
expectation of an ever-reachable connectivity. It is here where virtualization
plays an important role. In fact, the re-design of some network components today
existing via hardware can transform a monolithic networking approach to a 
modular one, exploiting the flexibility that Virtual Network Functions 
(VNFs)\cite{moens2014vnf} can offer thanks to virtualization, closing the gap to
the use-case fulfillment defined by the NGMN alliance~\cite{iwamura2015ngmn}
which require a decreased time to set up and deploy network services
(specifically, from 90 hours to 90 minutes~\cite{networld20202014role}).

\subsection{5G architecture}

With the constraints placed on the requirements formulated by the NGMN alliance,
5G envisage a multi-layered architecture, based on three main
layers~\cite{alliance20155g}:
\begin{itemize}
\item \textbf{infrastructure resource layer}: physical resources that are 
exposed via a virtualized interface, and that can be monitored using specific 
APIs
\item \textbf{business enablement layer}: where a library of functions and
  deployment is contained, and its configuration is accessible via APIs
\item \textbf{business application layer}: layer that contains specific
  applications and services of the operator
\end{itemize}

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.55]{5garchitecture}
  \caption[5G Architecture]{5G Architecture~\cite{alliance20155g}}
  \label{chap:background:img:5garchitecture}
\end{figure}

This separation in layers allows to easily manage multiple identities 
differently: an Internet Service Provider (ISP) could manage a multitude of 
physical infrastructures and have multiple business enablements dislocated 
along an entire continent for example, but it could decide to have only a 
single centralized business application deployment that manage all the other 
layers resources.

\subsubsection{Network slicing}
The role of a \emph{network slice} in a 5G architecture is to specifically
handle the Control-plane of a particular service (e.g. smartphones traffic,
autonomous driving, massive IoT), deploying resources in a manner that assure
the required latency, security and reliability. While some very peculiar and
legacy service could require specific hardware, the common resources between
services could be shared in a virtualized way, providing auto-scaling
capabilities in services that are under heavy network pressure.

\section{The VIBES project}
This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
requirement is to enhance the performance of end-to-end IP based that involves
satellite connections. To reach this goal, the project specifications suggest to
exploit the 5G incoming technology and use the NFV-MANO architecture to perform
first packets elaboration and performance improvement and finally TCP/IP
satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
VIBES project proposed five technical requirements:
\begin{enumerate}
  \item Analysis of the applicability of current and new Internet protocols in
  the proposed VNF-PEP architecture
  \item Implementation of a VNF-PEP prototype
  \item Building of a PoC test platform
  \item VNF-PEP validation and performance tests on 5G use-cases
  \item Demonstration test-bed management
\end{enumerate}

\begin{figure}[t]
 \centering
 \includegraphics[scale=1]{vibes_logo}
 \caption{ViBeS project logo}
 \label{chap:background:img:vibes_logo}
\end{figure}

\subsection{VNF-PEP architecture and internet protocols}
VNF-PEP is the softwarization of the usual PEP technology. One of the goals of
the project is to research and include this technology in the usual VNF
ecosystem to include without change satellite communications to the SFC. The
architecture of the systems will be based on clusters of containers whose
management and orchestration will follows the ETSI MANO architecture and SDN
principles. Thanks to VNF-PEP will be possible to dynamically take decisions on
how to optimize a certain communication.

%The analysis of this topic revealed to be trivial: since internet has many
%different protocols that would become infeasible to support all of them at the
%same time, packet encapsulation present itself as the only feasible solution:
%every packet incoming in the VNFs has already been encapsulated by a generic
%packet encapsulator/decapsulator,
%\todo{Also I think that the main purpose of UDP encapsulation is not to hide 
%the protocol used on the edges but 1) use a ``quick protocol to exchange data 
%among VNFs and 2) we are hiding the path not the protocol itself. As we were 
%discussing, we are creating some sort of proxy, so the aim will be the same 
%even if we support a plethora of protocols.''} making the whole architecture 
%independent from the protocol a particular flux of data. To achieve this, 
%several solutions have been studied, and at the end packet encapsulation with 
%TCP split (for TCP sessions) have been chosen. The rationale that guided us on 
%this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
%reference with the section of the explanation}
%
%\subsection{VNF-PEP prototype}
%
%Since the requirement for the whole system (MANO+PEP) were too challenging, the 
%goal shifted into creating a MANO test-bed and a working NFVI, excluding PEP. 
%The VNF architecture was shaped following the container orchestrator we decided 
%to use. An more detailed architectural implementation can be found at 
%Chapter~\ref{chap:archimpl}.
%
%\vspace{0.5cm}
%
%Starting with the first step of creating a MANO able to process incoming data
%packets through VNF functions, we encountered that many networking tools already
%present in the market required some tweaking and some integration, shifting our
%goal to create a complete European Telecommunications Standards Institutes
%(ETSI) Management and orchestrator (MANO) test-bed instead, following the
%specifications suggested in the RFC 7665, thus implementing only the first three
%requisites, without digging in the satellite data flow optimization. In
%particular, we discovered how, these tools, were suitable to create ETSI MANO
%and VNFs using virtual machine or exploiting cloud technologies, while they were
%not designed with enough flexibility an integration with Docker. In the next
%chapter we are going to take a deep analysis of the cited technologies 
%(Section~\ref{chap:prjan:sec:tech}, in order to provide to the reader enough 
%context to be able to understand our design choices we are going to describe 
%later.
%
%\noindent With this in mind, we performed a requirements analysis described in
%the next chapter.\todo{This should be put at the end of the introduction.}
%