### 2018-11-22 res/sections/11-ProjectScheduling.tex
Add schedule analysis and related problems

\section{Project Scheduling}
\label{chap:prjan:sec:prjsche}

Before starting this project a time requirements has been carried out, to see
its feasibility. We calculated at least six month necessary to be able to create
a working proof of concept (PoC) that allowed us to perform tests and, in
general, to have a feedback about the solution.

\subsection{Scheduled timeline}
Since our knowledge about the technologies was very poor at the beginning, we
preferred to invest a good chunk of time on them, to understand the already
present products available in the market and to choose what we really needed.
Additionally, at the end of this phase we though it was necessary to try these
products to see it they really fitted our necessities. After this phase, we
planned to actually starting implementing the missing components to create and
to fulfill the requirements defined in Section~\ref{chap:prjan:sec:req}. Since
the required technologies, Java and C/C++ are known to the authors, we estimated
an eight-week period related to coding and software implementation. After that,
a two weeks were planned in order to perform tests and gather metrics about the
system built, and too see how much the job done satisfied our expectations.
Finally, we estimated four weeks to group and write down all the information and
considerations we got during the thesis.

\subsection{Actual timeline}
Unfortunately, many difficulties slowed our project development, starting from
the beginning. The total time required to finish all the requirements decided
was of 37 weeks, namely nine months circa. Many problems emerged during the
study of the technologies, that revealed to be more complicated than expected,
especially to deploy and play with. Time was lost to obtain resources from the
University that was not initially planned. On top of that, we encountered little
bugs we had to manually fix, like for example in the Floodlight
Controller\footnote{See
  \url{https://github.com/floodlight/floodlight-webui/pull/10}}, that
additionally slowed our technologies exploration. This not-planned digging in
other projects' code cost us time and resources, delaying this phase of eight
weeks, for a total of 17 weeks. The software implementation revealed to be
irksome too. We end up underestimating the time required for this phase, because
we did not consider all the issues related to work with system calls. Moreover,
we had to implement more components than planned, in order to be able to
actually transmit data between two senders. Python, another program language,
had to be used to create proxy in order to carry out proper packet
encapsulation. This language, however simple, required some time to be picked up
by us, resulting in additional delay. In a nutshell, the additional time
required was of eight weeks, for a total of 16 weeks. Finally, the time to write
down the following thesis was limited due, and we had to write it down in a
total of two weeks, instead of the four initially planned.