### 2018-11-24 res/sections/07-Introduction.tex
New intro

\chapter{Introduction}
\label{chap:intro}

Through the history of Internet, many factors changed and new innovations came,
but at the end of the day a fundamental key remains: connection performances
are, indeed, what the users really want.
Today, connectivity plays an important role in everyone's life. Several decades 
brought solutions with enhanced wired and wireless connectivity, like 
Ethernet, Fast Ethernet, 2G, 3G, LTE, and better communication protocols such 
as a whole plethora of TCP standards in order to keep up with the continuous 
surging of network traffic. Internet Service Providers to be able to 
efficiently process the whole quantity of packets passing through the network, 
applies several optimizations along the packet path. Until this day, these 
optimizations have mostly been based on costly, in hardware solutions. In 
particular, these Network Functions are able to perform only generic packet 
elaboration, since they do not know the Quality of Service required by every 
connection. It is in this context that 5G proposes a new way to approach 
networking, suggesting the virtualization of this Network Functions, making 
their deployments faster and cheaper. On top of that, it proposes specific 
services for determined kind of data. If in the past a train running a 
300km/h had to use the same network services of an IoT device sending data 
updates every hour, now 5G proposes two different and efficient solutions for 
this completely different services. This separation is performed through the 
creation of determined network components called Network Slices. This new 
possibility of offering the network with specialized class of services, 
combined with emerging virtualization technologies like cloud computing, 
hypervisor-virtualized machine or container-based solutions enables the 
deployment of Virtualized Network Functions within seconds running on generic 
purpose or custom hardware. While there may be some performances degreadations, 
it has been demonstrated that their time-to-deploy and the possibility of 
automation they offers are unprecedented~\cite{nguyen2017sdn}. On top of 
that, specialized software acts as a MANager and Orchestrator (MANO) for this 
VNFs, regulating their lifecycle. While most of the current proposals are based 
on hypervisor on cloud solutions, in this thesis we are going to see a 
proof-of-concept of a part of a Network Slice, equipped with a custom MANO 
solution, based on the new container-based technology, called Docker, on a 
recent orchestrator framework: Kubernetes.

\section*{Document organization}

This thesis is organized as follows:
\begin{itemize}
 \item Chapter~\ref{chap:background} describes background information about 5G 
and the technologies used to build an ETSI compliant architecture
 \item Chapter~\ref{chap:rel_wk} gives an insight of correlated works on this 
field
 \item Chapter~\ref{chap:prjan} explains the analysis performed on the 
requirements provided and the design choices performed.
 \item Chapter~\ref{chap:archimpl} provides an in-depth view of the 
high-level architectural implementation.
 \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.
\end{itemize}