### 2018-11-27 res/sections/10-ProjectAnalysis.tex

\chapter{Project Analysis}
\label{chap:prjan}

As in chapter~\ref{chap:intro}, the creation of a full ETSI MANO compliant
architecture, following the suggestions described in RFC 7665, require the
integration of multiple tools (some of them have been already briefly introduced
in~\ref{chap:intro}). The tools were chosen after their analysis, where pro and
cons have been studied and compared to the necessity of this thesis. Since the
project has been developed with the cloud provided by the University of Padua,
Openstack (which a product analysis can be found later
in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.

\subsubsection{Openstack}
\label{chap:prjan:sec:openstack}
Created in 2010 by a joint operation between Rackspace Hosting and
NASA~\cite{openstackWebsite}, Openstack allows on-premise cloud installations,
using bare-metal resources to provide common cloud services as object storage,
virtual machine deploy, virtual networking. Furthermore its modularity offers
the possibility to add additional components, even proprietary, to achieve a
complete cloud solution.
\begin{figure}[t]
 \centering \includegraphics[scale=0.58]{openstack_logo}
 \caption{Openstack logo}
 \label{chap:prjan:img:openstack_logo}
\end{figure}

\subsubsection{Kubernetes}

As already introduced before, Kubernetes is an open source software framework
specialized in container management and orchestration. Developed by Google in
2014 and written in Go~\cite{k8sGit}, it allows different machines (called nodes
from now on) to create an abstraction layer and to form a cluster where is
possible to run Docker containers without bothering about hardware constraints.
Nodes can have specific roles and specialization, all of them managed by one or
many master nodes, that actually manage the orchestration. On top of that,
Kubernetes is virtualization-agnostic, meaning that it offers the possibility to
change the virtualization software in use (i.e. from Container-based technology
to Virtual machine software).
\begin{figure}[h]
 \centering \includegraphics[scale=0.35]{kubernetes_logo}
 \caption{Kubernetes Logo}
 \label{chap:intro:img:k8s_logo}
\end{figure}

\subsubsection{Docker}

In recent years, with the new hardware capabilities and the recent development
of in-kernel virtualization systems (such as Hyper-V) this technology begin to
be adopted. Virtualization allows to run different systems on the same machine,
making them completely isolated and then more resilient to failures. The back of
the medal, though, is that virtual machines require large amount of resources,
especially memory, because solutions like copy-and-write of in-memory pages are
not viable anymore (since the kernel gets duplicated too), leading to
duplication of loaded libraries and assets in the memory of the host. For large
deployment containing only simple services (e.g. a back-end application serving
a website or a database) this ends up in a waste of resources. \todo{Talk about
  lxc containers and how Docker solved magically the problem.} It is here that,
in 2013~\cite{dockerWebsite}, Docker was created, basing its solution on an
already existing product: Linux Container (LXC). From this framework, Docker
built an entire ecosystem, consisting in a client/server model giving the
possibility for users to simply launch, scale and delete containers (locally or
remotely with Docker-machine), a repository system where images, a layer-based
``core'' where containers start running from when they are launched, can be
stored and tagged in a sort of version control system. Finally, another couple
of solutions, called Docker-compose, and Docker Swarm have the purpose to,
respectively, orchestrate containers in a single or a clustered system.
\begin{figure}[t]
 \centering \includegraphics[scale=0.7]{docker_logo}
 \caption{Official Docker logo}
 \label{chap:intro:img:docker_logo}
\end{figure}

\vspace{0.5cm}

% The development of the ETSI MANO test-bed involved the use of several 
% tecnhologies already in the market, and in some cases the creation of ad-hoc 
% solutions to solve particular issues. Here, a brief listing of the 
% well-estabilished technologies used is performed, describing even solutions 
% that at the end weren't chosen because they didn't fit with the defined goals.

\subsection{MANO}

Another fundamental component in the architecture is the MANO, that has to 
manage the Network Services and the deployment of new VNFs. In coordination 
with Kubernetes, it composes the core of the test-bed architecture. Initially 
identified with Openbaton (the component is described later 
in~\ref{chap:prjan:sec:tech:sub:other:sub:openbaton}), afterword it has been 
discovered to be not flexible enough to support other kinds of virtualization 
techniques than pure or in-cloud virtual machines\todo{Rewrite this}. 
Since this was a strong project requisite that could not be easily changed, it 
has been decided to opt for the creation of a smaller and simpler clone that 
offered a full integration with Docker and Kubernetes.

\subsection{SFC}
The tools analysis revealed little about already existent VNF or NS
implementations, so it was necessary to fill the gap. Since this one of our
thesis goals, we developed a complete solution to create a complete SFCs, trying
to standardize our vocabulary as much as possible, and giving basic
documentation when possible.

\subsection{Technologies used during development}

Since the test-bed is composed by different components and thanks to its 
modularity, the implementation of the single components has been done with the 
help of additional tools, illustrated below

\subsubsection{Docker-compose} 
Docker-compose is a tool that allows services orchestration. It automates most 
of the tasks that should be performed by hand when launching one or multiple 
docker containers. It's particularly useful when more containers have to 
interact 
together, because\todo{Write down docker-compose functionalities and describe 
the product a little bit} all the services rules are described in a simple 
YAML file. With the birth of Docker Swarm, work has started to port this tool 
to 
the Swarm framework. At the time of this writing, the procedure to port a 
Docker-compose configuration in Docker Swarm (which a description of the 
product can be found in 
sub-section~\ref{chap:prjan:sec:tech:sub:other:sub:swarm}) is still not 
completely transparent, and it requires a sort of compilation, where 
Docker-compose bundles all the necessaries resources into a binary file that 
Docker Swarm will use to allocate the necessary resources.

\subsection{Other technologies evaluated}

Additional components have been evaluated during the development of the 
test-bed, and here we propose a list of the most important ones. Some of them 
where discarded as no more necessary, or during the evolution of the project 
itself they have become obsolete.

\subsubsection{Docker Swarm}
\label{chap:prjan:sec:tech:sub:other:sub:swarm}
Introduced in 2014~\cite{swarmGit}, Docker Swarm allows multiple Docker nodes to
cluster together and be seen as one logical unit. This layer completely makes
the underlying infrastructure transparent: data management, as network one, are
totally managed by Docker. Recently, Docker-compose support has been introduced,
making this one-node tool available for clustered systems. At the end of the
day, on one hand Docker Swarm provides an easy way to set up a cluster system
with all the tools configured out-of-the-box, without the necessity to set
networking configurations or installing ad-hoc storage solutions. On the other
hand, though, it doesn't have all the customization options that Kubernetes
offers, thus making the product less flexible. This fundamental characteristic,
although, makes the two solutions have different use cases leading to different
market needs.

\subsubsection{OpenVSwitch}
Started in 2009~\cite{ovsGit}, the OpenVSwitch project aims to create an
open-source multi-layer virtual switch. It provides a switching stack for
hardware virtualization environments, implementing several interfaces and
protocols. It has also the possibility to run through Hypervisor virtualization
but it can run under Docker too~\cite{ovsDocker}. OpenVSwitch can route packets
from incoming connections following rules, that can be ``pushed'' in its routing
table from software defined as \textit{controllers}. Additionally, it supports
the OpenFlow communication protocol, in order to set previously cited routing
definitions.

\subsubsection{Floodlight}
\label{chap:prjan:sec:tech:sub:other:sub:floodlight}
Community-developed and open source, Floodlight is an OpenFlow Controller
written in Java at the end of 2011~\cite{floodlightGit}. It allows various
applications to run on top of it. Floodlight has the role of controlling and
inquiring an OpenFlow network, querying the applications integrated with the
purpose of solving different user needs over it. When instantiated, OpenVSwitch
routers have to register to the controller in order to gain routing rules.

\subsubsection{Openbaton}
\label{chap:prjan:sec:tech:sub:other:sub:openbaton}
Patronized by the Institute Fraunhofer, it is an open-source, customizable NFV
MANO-compliant framework, that offers many configurations options, and since
it's written in Java, there is the possibility to add components (plug-ins)
dynamically. Openbaton can be installed as a normal computer program (they offer
packages for the most common distros) but it can be deployed on Docker
containers too. It's designed to work with cloud providers like Amazon or Google
Compute engine, but it offers compatibility with on-premise cloud solutions like
Openstack, where it deploys virtual machines using the API the platform offers.
It can store VNF configurations saved as Topology and Orchestration
Specification for Cloud Applications (TOSCA) Yet Another Markup Language (YAML)
and it can handle multiple Point of Presence (PoP). It offers a VNF lifecycle
management out-of-the-box, with the possibility to customize it based on the
user needs. Its modular design allows developers to change parts of the codebase
with custom ones, making the product really flexible to different platforms.
Performing this operation requires a deep knowledge of the framework though.
\begin{figure}[h]
 \centering \includegraphics[scale=0.45]{openbaton_logo}
 \caption{Openbaton logo}
 \label{chap:prjan:img:openbaton_logo}
\end{figure}