### 2018-11-27 res/sections/20-VNFImplementaton.tex

\chapter{VNF and Network Service implementation} 
\label{chap:vnf_ns_impl}

%Even though SFCs and VNFs are part of the ETSI architecture, the complexity
%behind was so high to require particular care and a separated study. Indeed,
%the SFCs are the core of the whole project: packet transmission and elaboration
%is made here. Challenges regarding this topic are building a system that is
%flexible enough to allow dynamic packet routing, support scalability and
%provide methods to recovery from failures. With this challenges packet
%transmission revealed to be a fundamental point to choose carefully.