### 2018-11-27 res/sections/03-Inscription.tex
Add acknowledgement and first citation

%**************************************************************
% Dedica
%**************************************************************
\cleardoublepage
\phantomsection
\thispagestyle{empty}
\pdfbookmark{Inscription}{Inscription}

\vspace*{3cm}

\begin{flushright}{
	\slshape
	Do not shorten the morning by getting up late, or waste it in unworthy occupations or in talk; look upon it as the quintessence of life, as to a certain extent sacred. Evening is like old age: we are languid, talkative, silly. Each day is a little life: every waking and rising a little birth, every fresh morning a little youth, every going to rest and sleep a little death.} \\
	\medskip
  Arthur Schopenhauer
\end{flushright}

\medskip