### 2018-11-27 res/sections/04-Credits.tex

%**************************************************************
% Ringraziamenti
%**************************************************************
\cleardoublepage
\phantomsection
\thispagestyle{empty}
\pdfbookmark{Credits}{credits}

%\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

%\section*{Credits}

\noindent \textit{Acknowledgements}\\
\bigskip

\noindent When I began my journey at University, five years ago, I was only
a boy. A boy curious about IT and geeky stuff, with the humble dream of becoming
proficient with computing programming. During the years, the dream gradually
became a goal to reach, a destination. The goal changed, becoming more
ambitious, and now I am finally going to reach it\dots But I would not be able
to make it without the multitude of people that I luckly have around me. It is
for the power of this people if today I am here writing these acknowledgements.

\vspace{0.3cm}

\noindent A big thank for my friends, whose power of cheering me up and make
me smile saved myself different times from the risk of failing this journey.
Thanks because you where always here when I needed you, and you always found the
right words to say, or the right thing to do. Big part of my success is thanks
to them, who have pushed me in places I would not be able to reach alone. In
particular I would like to thank Giovanni, Giorgio, Michele, Gianluca and
Umberto for the good times we end up spending every weekend. On top of that, I
would like to say also thanks to Marco, Federico, Matteo, Mirko, Emanuele,
Stefano, Sara, Lisa and Eduard that closely witnessed with me the good and the
bad moments of the last two years of this journey. The time spent together with
you guys will not be easily forgotten.

\vspace{0.3cm}

\noindent Last but not least, I would like to praise my best friend, Franco,
which has the power to keep me rolling in every aspect of my life, even when I
am without energy to keep up or when I am completely lost, and which is alway
with me, regardless of the physical distance that separate us. Thanks.

\vspace{0.3cm}

\noindent Not only friends helped me to complete my journey. Additional
acknowledgements go to my favourite sport, Judo, that allowed me to mentally e
spiritually grow and to become what I am now. It is for people that never gave
up like you, Pierluigi, if I am now able to keep a cool head even in the worst
of the storms.

\vspace{0.3cm}

\noindent Other thanks go to my parents, which supports my every day and always
believed in me, no matter what.

\vspace{0.3cm}

\noindent Finally, thanks to my supervisor, Prof. Armir Bujari, to have
mentored me during the design and development of this thesis.

\vspace{0.5cm}

\noindent\textit{\myLocation, \myTime}
\hfill \myName

\endgroup