### 2018-11-28 res/sections/40-FutureWork.tex
Finish future work draft

\chapter{Future Work}
\label{chap:fw}

The realization of this architecture proposal is not finished yet. Indeed, a lot
of functionalities could be implemented and new components coul be added. We are
now going to see possible future additions to improve the system.

\section{Security concerns}

One of the aspects that is usually not really given the required attention is
the one regarding security. In this context, the security we talk is the one
about the whole system, and not the single VNFs implementations. Fortunatly,
Kubernetes is equipped with out-of-the-box access management options. A good way
to manage user authentication in the system should be integrate the Role Based
Access model provided by Kubernetes directly in Harbor. In fact, Harbor does not
come with security features, and the exposed APIs are usable by everyone. This
could lead to possible Denial of Service (DoS) attack of the disruption of NS
services by a malicious attacker.

Another security aspect which is important to provide is the isolation between
the network slices. This isolation, in our implementation, should be achieved
clearly identifing sets of nodes (the different slices) and restricting their
possibility to talk and interact with the other sets. This kind of restrinction
should be achieved using different group Policy through Kubernetes.

\section{Kubernetes federation}

A single Kubernetes cluster presents limitations in terms of number of nodes
available on a single cluster and regarding the geographical position of this
nodes. Indeed, nodes in a cluster should be located or in the same network or
geographically closer to each other, in order to ensure show latences between
data exchange and Pods communications. An available solution, introduced in
version 1.9 of Kubernetes, is the possibility to create federated Kubernetes
cluster. This solution ensures a proper division between groups of nodes, and
the possibility to ensure better performances creating cluster with nodes closer
to each other. Unfortunatly, we werent able to create a proper Kubernetes
federation. On of the reason was related the current immaturity of the project,
that could have created and instable cluster federation.

\section{Dynamic chain update}

The importance of having NS services updated as smooth as possible is without 
doubts a feature that would ensure a continue service availability avoiding its 
disruption. This kind of approach, though, requires particular attention about 
keeping the current connections alive. This means at least keeping the old 
chain up while deploying the new one, and gradefully redirect all the 
new traffic from the chain to replace, until none is left. Only after this 
operation, the old chain definition can be completely decomissioned and the 
related VNFs removed. In our current implementation, the NS update 
functionality relies on a simpler approach: when an update gets ordered, 
Kubernetes gets instructed to remove the old chain and deploy the new one. 
Meanwhile, all the Roulette controllers get updated with the new NS route. The 
back of the medal of this soulution is, while it keeps the sistem implementation 
simpler, that packets current travelling along the old NS could be lost during 
the chain updating.