### 2018-11-28 res/sections/20-VNFImplementaton.tex
Started sfc implementation

\chapter{Service Function Chain Implementation} 
\label{chap:vnf_ns_impl}

The complexity behind the creation of a working Service Function Chain system
was so high to require particular care and a separated study. Indeed, the SFCs
are the at the base of the whole architecture: packet transmission and
elaboration is made here. Challenges regarding this topic are building a system
that is flexible enough to allow dynamic packet routing, support scalability and
provide methods to recovery from failures. In this chapter, a brief explanation
about the challenges and the issues encountered building this system will be
given.

\section{Harbor}

As already cited, Harbor acts as a MANO, orchestrating and managing VNFs and the
related services. From the SFC point of view, it carries consistency checks when
a NS definition is created. The checks are carried out based on the VNF ID.
While the VNF definitions corresponds to a YAML Kubernetes configuration, the NS
one is a JSON array, for example:

\begin{lstlisting}[caption={Example of SFC definition}, captionpos=b,
                   language=json, label=chap:impl:lst:sfcexample]
{
    "ns" : [
                {
                    "id": "astaire-service"
                },
                ...
            ]
}
\end{lstlisting}

\section{Kubernetes Pods}

All the SFC-related components have been developed with the idea of being
containerized. This is a foundamental concept, because in this way all the
components have the possibility to run inside Kubernetes, granting them the
scalability and persistence benefits. On top of that, the use of
\verb!Deployment! granted the possibility to perform easy an update of already
existing components.

\section{Approved proposal}

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{finalreview}
  \caption{Final review schema}
  \label{chap:impl:img:finalreview}
\end{figure}

The approved proposal for the creation of a proper chain has been based on four components:
\begin{itemize}
\item Astaire;
\item Harbor;
\item Ironhide;
\item Roulette.
\end{itemize}