### 2018-11-29 res/sections/05-Abstract.tex
Fix margins and section title

%**************************************************************
% Abstract
%**************************************************************
\cleardoublepage
\phantomsection

\thispagestyle{empty}

\vspace*{\fill}
\par
\begingroup
\leftskip4em
\rightskip\leftskip
\section*{\centering Abstract}
The importance of Internet is growing every year. Mobile and wired traffic are
continuously surging, asking for better performances. On this day, ISPs rely on
optimized, ad-hoc and often in-hardware nework solutions. These network
functions are costly and require time to be deployed and mantained. In this
context, upcoming virtualization technolgies like hypervisor-based and
container-based are providing the possibility to create virtualized network
functions, that have the ability to be up and running in a matter of seconds.
VNFs are the managed and orchestrated by a specific application, the MANO that
handles their lifecycle.

In this thesis we are going to propose an Infrastructure Management and
Orchestrator plane Proof-of-Concept, based on the container virtualization
technology and on the 3GPP Release 15 specifications. In the current PoC, Docker
and the container orchestrator Kubernetes play a fundamntal role.
\par
\endgroup
\vspace*{\fill}
%\bigskip