### 2018-11-29 res/sections/03-Inscription.tex
80 columns

%**************************************************************
% Dedica
%**************************************************************
\cleardoublepage
\phantomsection
\thispagestyle{empty}
\pdfbookmark{Inscription}{Inscription}

\vspace*{3cm}

\begin{flushright}{
	\slshape Do not shorten the morning by getting up late, or waste it in
	unworthy occupations or in talk; look upon it as the quintessence of life, as
	to a certain extent sacred. Evening is like old age: we are languid,
	talkative, silly. Each day is a little life: every waking and rising a little
	birth, every fresh morning a little youth, every going to rest and sleep a
	little death.} \\ \medskip Arthur Schopenhauer
\end{flushright}

\medskip