### 2018-11-30 res/sections/12-Requirements.tex

\section{Requirements}\label{chap:prjan:sec:req}
In the following section are described the main requirements of this project.

\begin{longtable}[c]{p{0.3\textwidth}p{0.7\textwidth}}
\hline
\multicolumn{1}{c}{\textbf{Requirement}}                                & \multicolumn{1}{c}{\textbf{Description}}                                                                                                                                                       \\ \hline
\endhead
%
\hline
\endfoot
%
\endlastfoot
%
1. Automatic NFVI creation and configuration                            & It must be possible to create the environment in which the SFC will be deployed in an automatic manner. The system must support Docker container and must be deployed on a clustered environment. \\
2. Implementation of chain orchestration and management tools           & The final work must be comprehensive of a minimal software implementation that allows the deployment, the management and the orchestration of VNFs and their lifecycle.                      \\
3. Automatic configuration of chain orchestration and management tools & Software instrumentation for VNFs and SFCs management must be configured automatically, at startup.                                                                                              \\
4. Developing software for manual SFC deployment                        & An instrumentation to deploy a chain of containers in which VNFs can run must be developed. These containers must be able to communicate with each other.                                                \\
5. VNF fault management                                                 & The system must be able to recover a faulty container automatically.                                                                                                                      \\
6. VNF scalability                                                      & The system must be able to scale the number of containers implementing a function, both augmenting and decrementing their replicas.                                                               \\
7. SFC definition saving                                                & The system must be able to save a definition of a chain.                                                                                                                                        \\
8. Chain definition deploy/stop/update                    & The system must be able to allow the deployment of a chain from a saved definition, stop a running SFC and update it.                                                                          \\ \hline
\caption{Table of requirements}
\label{chap:prjan:tab:req}\\
\end{longtable}