### 2018-12-05 thesis/res/sections/11-ProjectScheduling.tex
Presentation integration (#18)
Integrate presentation in thesis repository

\section{Project Scheduling}
\label{chap:prjan:sec:prjsche}

Before starting this project a time scheduling has been carried out, to see
its feasibility. We calculated at least six months necessary to be able to create
a working proof of concept (PoC) that allowed us to perform tests and, in
general, to have a feedback about the solution.

\subsection{Scheduled timeline}
Since our knowledge about the technologies was very poor at the beginning, we
preferred to invest a good chunk of time on studying them, to understand the
already present products available in the market and to choose what we really
needed. Additionally, at the end of this phase we though it was necessary to try
these products to see if they really fitted our necessities. After this phase,
we planned to actually starting implementing the missing components to create
and to fulfill the requirements defined in Section~\ref{chap:prjan:sec:req}.
Since the required technologies, Java and C/C++ are known to the authors, we
estimated an eight-week period related to coding and software implementation.
After that, two weeks were planned in order to perform tests and gather metrics
about the system built, and too see how much the job done satisfied our
expectations. Finally, we estimated four weeks to group and write down all the
information and considerations we got during the thesis.

\subsection{Actual timeline}
Unfortunately, many difficulties slowed our project development, starting from
the beginning. The total time required to finish all the planned requirements
was of 37 weeks, namely nine months circa. Many problems emerged during the
study of the technologies, that revealed to be more complicated than expected,
especially since some frameworks were more intricate to deploy and to try them
out. Time was lost to obtain resources from the University that was not
initially planned. On top of that, we encountered little bugs we had to manually
fix, like for example in the Floodlight Controller\footnote{See
  \url{https://github.com/floodlight/floodlight-webui/pull/10}}, that
additionally slowed our technologies exploration. This not-planned digging in
other projects' code cost us time and resources, delaying this phase of eight
weeks, for a total of 17 weeks. The software implementation revealed to be
irksome too. We end up underestimating the time required for this phase, because
we did not consider all the issues related to work with system calls. Moreover,
we had to implement more components than planned, in order to be able to
actually transmit data between two senders. In a nutshell, the additional time
required was of eight weeks, for a total of 16 weeks. Finally, the time to write
down the following thesis was limited too, and we had to write it down in a
total of two weeks, instead of the four initially planned.